### 2022-02-14 introduction/introduction.tex
more

Proof assistants (or interactive theorem provers) are programs which allow users
to describe a theory in a formal way and to prove the statements about this
theory using rules of the inference system the proof assistant is based on.
These programs can then check if the given proof is correct. As their core is usually
small and well-tested, proofs that pass the checking phase are considered trustworthy.

>>>>

Proof assistants (or interactive theorem provers) are programs which allow users
to describe a theory formally and to prove the statements about this theory
using rules of the inference system the proof assistant is based on. These
programs can then check if the given proof is correct. As their core is usually
small and well-tested, proofs that pass the checking phase are considered
trustworthy.

They proved indispensible in proving very complex mathematical theorems in a
trustworthy manner. For example, four-color theorem \cite{gg-2008-four-color}
and Kepler conjecture \cite{th-2015-kepler} were proven inside proof assistants,
ending the need to trust hundreds of pages of hard mathematical proofs. Proof
assistants are also used to show correctness of complex software such as
compilers \cite{xl-09-compcert} or operating systems kernels \cite{gk-09-sel4}.
This verified software can be used in safety-critical scenarios such as nuclear
plants or autonomous vehicles.

The proofs need to be spelled out in minute details to be accepted by a proof assistant.
Even though many proof assistants offer some sort of automation, seemingly simple
statements cannot be automatically proved. This is one of the reasons why large
verification projects can take years to finish.

Integrating proof automation right into the proof assistant is not an easy task.
Most proof assistants are based on variants of higher-order logic, which is more
expressive than first-order logic, but in general does not even allow complete
proof calculi. To provide some level of automation, proof assistant developers
use the efficient first-order provers as follows. First, the proof goal is
translated from higher-order to first-order logic. Then the first-order prover
is run on the translated problem. If the first-order prover finds the proof, the
proof is reconstructed in the proof assistant. This approach is implemented in
modern assistants such as Coq \cite{bc-04-coq}, HOL Light \cite{jh-09-hol-light}
and Isabelle \cite{lc-88-isabelle} using \emph{hammers} CoqHammer \cite{ck-18-coqhammer}, HOL(y)Hammer \cite{ku-15-holyhammer}
and Sledgehammer \cite{bn-10-sh}, respectively.
